---
title: "covid 19"
author: "Daniella Sanchez"
date: "Summer 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
dd2 <- kRdata %>% group_by(sex,result) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```

The barcharts show that the make dealths related to covid make up to 60% of the total death and females make up to the other 40% of the death.

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Our numerical results show us that out of 100, males make up to 64% of covid related death to females at 36%. This also shows that males are twice as likely to die from covid than females. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Fisher's Exact Test for count data shows that males are 2.3 times more likely to die from COVID compared to females. Our p-value is also a very small number which allows me to reject the null hypothesis that states that COVID deaths are independent of sex. 

## Conclusions

Our null hypothesis stated that COVID related deaths were independent to sex. First we added a Bar chart in order to show were the amount of deaths per 100. From here it shows that roughly 60% of deaths out of 100 were male and females were 40%. Then we ran the numerical results in order to have a more precise number of COVID related death between males and females. this gave us that 64 males die to COVID compared to 36 females. this also meant that we could state that males were more likely to die at a higher rate due to covid compared to females. 
